---
output: github_document
title:  Chapter 10 Examples -  Processing Data with purrr
---

The  `tidyverse` package `purrr` provides a comprehensive set of functions that can be used to iterate over data structure.

The idea of the `map()` function is to provide a mechanism to iterate over an input list or a vector. The general format of this function is `map(.x, .f)`, where:

- `.x` is a list, or an atomic vector. If .x has named elements, the return value will preserve those names.
- `.f` can be a function, formula, or vector. 

```{r}
library(purrr)

o1 <- purrr::map(c(1,2,3,2),function(x)x^2)
str(o1)
```

A *formula* (which is defined by `~`) can be used as a function argument too.

```{r}
o2 <- purrr::map(c(1,2,3,2),~.x^2)
str(o2)
o3 <- purrr::map(c(1,2,3,2),~.^2)
str(o3)
```

The package `purrr` provides a set of additional functions that specify the result type (e.g. double, character)

```{r}
library(dplyr)
library(purrr)
mtcars %>%
  dplyr::select(mpg,cyl,disp) %>%
  purrr::map_dbl(mean)
```


```{r}
library(repurrrsive)
library(purrr)
sw_films %>%
  purrr::map_chr(~.x$director) %>%
  unique()
```


You can also iterate over two input vectors using `map2()`

```{r}
means <- c(10,20,30)
sds   <- c(2,4,7)

purrr::map2(means,sds,~rnorm(5,.x,.y)) %>% str()
```

The function `group_split()` can be used to generate a list, which then can be processed with `map()`. Here we perform correlation analysis between windspeed and mean sea level pressure (we would expect a negative relationship).

```{r,warning=FALSE,message=FALSE}
library(aimsir17)
d_data <- observations %>%
           dplyr::filter(complete.cases(observations)) %>%
           dplyr::group_by(station,month,day) %>%
           dplyr::summarize(MaxWdsp=max(wdsp,na.rm=TRUE),
                            DailyAverageMSL=mean(msl,na.rm=TRUE)) %>%
           dplyr::ungroup()
d_data
```

```{r}
cor7 <- d_data %>%
         dplyr::group_by(station) %>%
         dplyr::group_split() %>%
         purrr::map_df(~{
                  corr <- cor(.x$MaxWdsp,.x$DailyAverageMSL)
                  tibble(Station=first(.x$station),
                         CorrCoeff=corr)
         }) %>%
         dplyr::arrange(CorrCoeff) %>%
         dplyr::slice(1:7)
```

```{r}
cor7
```
