---
output: github_document
title:  Chapter 6 Examples -  S3 Object System
---

You may have noticed that if you call `summary()` with two different types of data, you will get two different answers.

The results of a linear model.

```{r}
mod <- lm(mpg~disp,data=mtcars)
summary(mod)
```

Compared to the results from processing a data frame.

```{r}
d <- subset(mtcars,select=c("mpg","disp")) |> head()
summary(d)
```


An S3 object is a base type *with a defined class attribute*, and we can explore this by looking at the `class` attribute associated with a data frame.

```{r}
# Show the class attribute using attr
attr(mtcars,"class")
# Show the class attribute using class
class(mtcars)
```

An example of a generic function is `summary()`, and two (of the many) methods that belong to this generic function include `summary.lm()` and `summary.data.frame()`. Notice that the methods that belong to the `summary()` generic function have the format *summary.class_name(object, arg1, ...)*.

Here we show a simple example of defining a class and creating a new method based on the generic function `print()`.

```{r}
# define an atomic vector
v <- c(10.1, 31.2)
# Show v at the console
v
# Add a class attribute
class(v) <- "my_new_class"

# Implement a print method for this class, and load.
print.my_new_class <- function(o){
  cat("Hello World!\n")
}
# Show v again at the console
v
```
