---
output: github_document
title:  Chapter 8 Examples -  Data Transformation with dplyr
---

The package `dplyr` provides  functions to support data transformations, and also additional functions that can be used to summarize data.

First, we include the `dplyr` library, and the library `ggplot2`, which contains the tibble `mpg`.
```{r,message=FALSE}
library(dplyr)
library(ggplot2)
```
### Filtering rows with `filter()`

The function `filter()` is used to subset a tibble, in which all rows satisfying a condition are retained. This function conforms to the *architecture* of `dplyr`, in that it accepts a tibble, and arguments, and always returns a tibble. Note the use of the package name `dplyr::` in these function calls is optional.

```{r}
dplyr::filter(mpg,class=="2seater")

dplyr::filter(mpg,class=="2seater",hwy >= 25)


dplyr::filter(mpg, manufacturer == "lincoln" | 
                   manufacturer == "mercury")

dplyr::filter(mpg, manufacturer %in% c("lincoln","mercury"))
```


### Sorting rows with `arrange()`
The `arrange()` function orders the rows of a tibble by the values of selected columns, and these can be in ascending order (the default), or descending order, by using the function `desc()`. 

```{r,echo=T}
dplyr::arrange(mpg,cty) %>% slice(1:3)

dplyr::arrange(mpg,desc(cty)) %>% slice(1:3)
```

### Choosing columns with `select()`
The function `select()` allows you to subset columns from the input tibble using the column names. Note that `select()` will always return all of the observations for the selected columns.  

```{r,echo=T}
dplyr::select(mpg,manufacturer,model,year,displ,cty)

mpg %>% dplyr::select(manufacturer:year,cty)

mpg %>% dplyr::select(!(manufacturer:year))

mpg %>% dplyr::select(c(manufacturer:year,cty))

mpg %>% dplyr::select(starts_with("m"))

mpg %>% dplyr::select(contains("an"))

mpg %>% 
  dplyr::select(manufacturer:year,cty,hwy,everything())

```


### Adding columns with `mutate()`
The function `mutate()` adds new variables to a tibble, while keeping the original ones.

To explore `mutate()`, we take a reduced version of `mpg`, with six random observations and a smaller set of columns. Note that we use the dplyr function `sample_n()` to generate a sample from `mpg`, from two classes, "compact" and "midsize".

```{r,echo=T}
set.seed(100)
mpg_m <- mpg %>%
         dplyr::filter(class %in% c("compact","midsize")) %>%
         dplyr::select(manufacturer:year,cty,class) %>%
         dplyr::sample_n(6)
mpg_m
```


A vector of size 1 is added, and this is recycled to the correct length. 

```{r,echo=T}
mpg_m %>% dplyr::mutate(Test="A test")
```

A vector of the same length is added, and this vector is based on an existing column.

```{r,echo=T}
mpg_m %>% dplyr::mutate(cty_kmh=cty*1.6,
                        cty_2=cty_kmh/1.6)
```


### Summarizing observations with summarize()
The function `summarize()` creates a new tibble. It will have one (or more) rows for each combination of grouping variable.

```{r}
mpg %>% 
  dplyr::summarize(CtyAvr=mean(cty),
                   CtySD=sd(cty),
                   HwyAvr=mean(hwy),
                   HwySD=sd(hwy))
```

You can group by a variable, or a set of variables. This allows for summaries by group, which is very useful.

```{r}
mpg %>% 
  dplyr::group_by(class) %>%
  dplyr::summarize(CtyAvr=mean(cty),
                   CtySD=sd(cty),
                   HwyAvr=mean(hwy),
                   HwySD=sd(hwy),
                   N=dplyr::n()) %>%
  ungroup()
```

