---
output: github_document
title:  Chapter 5 Examples -  Matrices and Data Frames
---

### Matrices
In R, a matrix is a two-dimensional structure is created using the `matrix()` function, with the following arguments:

-   `data`, which are the initial values, contained in an atomic vector, supplied to the matrix,
-   `nrow`, the desired number of rows,
-   `ncol`, the desired number of columns,
-   `byrow`, a logical value (default is FALSE), that specifies what way to fill the matrix with data, either filled by row or by column,
-   `dimnames`, a list of length 2 giving row and column names respectively.

We now use this information to create a 3-by-3 matrix.

```{r}
set.seed(100)
data <- sample(1:9)
data
m1 <- matrix(data,
             nrow = 3, 
             ncol=3,
             dimnames = list(c("R1","R2","R3"),
                             c("C1","C2","C3")))
# Display the matrix
m1

# Show the number of rows and columns
nrow(m1)
ncol(m1)

# Show the matrix dimensions
dim(m1)
```

To subset a matrix, we supply an index for each dimension `[`row number, column number`]`, for example, here we extract the value contained in row 2, column 2.

```{r}
m1

# Extract the value in row 2, column 2
m1[2,2]
```

Other examples of subsetting include:

```{r}
# Extract first row and all columns
m1[1,]

# Extract first column, returned as a vector
m1[,1]
```

Matrix algebra operations are also supported:
```{r}
A <- matrix(1:4,
            nrow=2)
B <- matrix(1:4,
            nrow=2,
            byrow = T)
A

B
# Use matrix algebra to multiply two matrices
A%*%B
```

### Data Frames

On an intuitive level, a data frame is similar to a matrix, with a two-dimensional row and column structure, while on a technical level, a data frame is a list. Crucially, the elements (columns) of a data frame can be of different types, and this feature provides functionality for processing datasets as part of the operations research process. 

Here is a data frame:

```{r}
d <- data.frame(Number=1:5,
                Letter=LETTERS[1:5],
                Flag=c(T,F,T,F,T),
                stringsAsFactors = F)
d
summary(d)
```


