---
output: github_document
title:  Chapter 7 Examples -  Visualization with ggplot2
---

### Exploring relationships with a scatterplot
Creating a scatterplot to explore the relationship between two numeric variables is a convenient way to start our 
exploration of `ggplot2`

We include the library `ggplot2`, and will focus on the tibble `mpg`.

```{r}
library(ggplot2)
mpg
```

```{r, gg-fig-7-1, fig.cap = "Plotting displacement v city miles per gallon"}
ggplot(data=mpg, mapping=aes(x=displ,y=cty)) +
  geom_point()
```

We can set the `aes()` argument `color` to the particular variable we would like to color the plot by. 

```{r, gg-fig-7-2, fig.cap = "Coloring the points by the class variable"}
ggplot(data=mpg,mapping=aes(x=displ,y=cty,color=class))+
  geom_point()
```


Plots can be stored in a variable, and then added to using the `+` operator. Here the function `labs()` can be used to organise plot information.

```{r, gg-fig-7-4, fig.cap = "Configuring the plot using the labs() function"}
p1 <- ggplot(data=mpg,aes(x=displ,y=hwy,size=cyl,color=class))+
        geom_point()

p1 <- p1 + 
       labs(
         title = "Exploring automobile relationships",
         subtitle = "Displacement v Highway Miles Per Gallon",
         color = "Class of Car",
         size = "Cylinder Size",
         caption = "Sample chart using the lab() function",
         tag = "Plot #1",
         x = "Displacement (Litres)",
         y = "Highway Miles Per Gallon"
       )

p1

```


### Subplots with facets
Plots can be divided into subplots using `facet_wrap()`.

```{r,gg-fig-7-6, fig.cap = "Using facets to sub-divide plots"}
ggplot(data=mpg,aes(x=displ,y=cty))+
  geom_point()+
  facet_wrap(~class)
```

A 2D grid can also be arranged for subplots. Here the rows are the variable `class`, while the columns are the variable `drv`.

```{r,gg-fig-7-7, fig.cap = "Generating a grid of plots"}
ggplot(data=mpg,mapping = aes(x=displ,y=cty))+
  geom_point()+
  facet_grid(class~drv)
```


### Statistical transformations
Here we show a bar chart.

```{r,,gg-fig-7-8, fig.cap = "A bar chart show frequency counts of diamond cuts"}
ggplot(data=diamonds,mapping=aes(x=cut))+
  geom_bar()
```

And now, we draw a histogram using `geom_histogram()`

```{r,gg-fig-7-13, fig.cap = "Display the price histogram for all diamonds"}
ggplot(data=diamonds,mapping=aes(x=price))+
  geom_histogram(bins = 15)
```

Boxplots are useful representations of continuous data.

```{r,gg-fig-7-16, fig.cap = "Highlighting continuous variables using a boxplot"}
ggplot(data=mpg,mapping=aes(y=cty,x=class))+
  geom_boxplot()
```


### Covariation with `ggpairs()`

It is useful to visualize associations between variables.

```{r,gg-fig-7-17, fig.cap = "Visualising possible associations between three variables",message=FALSE}
library(GGally)
my_vars <- subset(mpg,select=c(cty,hwy,displ))
ggpairs(my_vars)
```

### Themes with `ggplot2`
We can use different themese when displaying plots.

```{r}
p1 <- ggplot(mpg,aes(x=displ,y=cty,colour=class)) + geom_point()

p1 + theme_bw() + ggtitle("theme_bw()")
p1 + theme_light() + ggtitle("theme_light()")
p1 + theme_dark()+ ggtitle("theme_dark()")
p1 + theme_minimal()+ ggtitle("theme_minimal()")
p1 + theme_classic() + ggtitle("theme_classic()")
p1 + theme_void() + ggtitle("theme_void()")
```
