---
output: github_document
title:  Chapter 4 Examples -  Functions, Functionals and the R Pipe
---

### Functions
The general form of a function in R is:

`function (arguments) expression`

where:

* `arguments` provides the arguments (inputs) to a function, and are separated by commas,
* `expression` is any legal R expression contained within the function body, and is usually enclosed in curly braces (when there is more than one expression),
* the last evaluated expression is returned by the function, although the function `return()` can be also used to return values.

Here is a function that returns only the even values in a vector.

```{r}
evens <- function(v){
  v[v%%2==0]
}

x1 <- 1:7

evens(x1)
```

A second function takes in a vector of random numbers, and remove any duplicates. To remove the duplicates, we will make use of the R function `duplicated()`, which returns a logical vector that contains `TRUE` if a value is duplicated. 

```{r}
my_unique <- function(x){
  # Use duplicated() to create a logical vector
  dup_logi <- duplicated(x)
  # Invert the logical vector so that unique values are set to TRUE
  unique_logi <- !dup_logi
  # Subset x to store those values are unique
  ans <- x[unique_logi]
  # Evaluate the variable ans so that it is returned
  ans
}
```

We can now call the function:
```{r}
set.seed(100)
v <- sample(1:6,10,replace = T)
ans <- my_unique(v)
ans

```

### Functionals

Functionals can be used instead of loops to iterate over data and return a result. One of the most important functionals that can be used is `lapply(x, f)`, which:

* Accepts as input a list  `x` and a function `f`, 
* Returns as output a new list of exactly the same length as `x`, where each element in the new list is the result of applying the function `f` to the corresponding element of the input list `x`. 

Here is an example of using `lapply()`, it calculates the minimum value of each list element. Note the looping action of `lapply()`, as it iterates over each list element.

```{r}
l1 <- list(a=1:4,b=10:11,c=20:25)
res <- lapply(l1,min)
res
```
### R’s native pipe operator

The native pipe operator in R, represented by the symbol `|>`, allows you to chain a number of operations together, without having to assign intermediate variables. The general format of the pipe operator is `LHS |> RHS`,  where `LHS` is the first argument of the function defined on the `RHS`.

```{r}
set.seed(200)
# Generate a vector  of random numbers
n1 <- runif(n = 10)
# Show the minimum the usual way
min(n1)
# Use the native pipe to generate the same answer
n1 |> min()
```


